---
title: "Applied Bayesian Statistics"
subtitle: "Class Assignment 2"
author: "Simon Roth \\\n nomis.roth@gmx.net"
date: "14.01.2019"
output: 
  pdf_document:
    number_sections: true
---


```{r setup, include=F}
knitr::opts_chunk$set(echo = T, fig.height = 5, fig.width = 7, fig.align = "center")
```


## Packages {-}

**Note** *This is Rmarkdown document which combines the entire code, outputs and text. If you don't have pacman installed, just get it once from CRAN and it will manage all the rest of the dependecies (for ever).*

```{r}
# install.packages("pacman")
pacman::p_load(
  tidyverse, rjags, purrr, tidyr, broom,
  texreg, ggthemes, janitor, knitr
)
ggplot2::theme_set(theme_bw())
set.seed(2018)
```

## Helper {-}

```{r, echo = F}
plot_jags_model <- function(model, terms = c("beta0", "beta1", "sigma")){
  
  tidy_bayes_dat <- model %>% 
    dplyr::tibble(output = .) %>% 
    dplyr::mutate(chain = as.character(1:n())) %>%
    dplyr::mutate(output = purrr::map(output, as_tibble)) %>% 
    tidyr::unnest() %>% 
    dplyr::group_by(chain) %>% 
    dplyr::mutate(step = 1:n()) %>% 
    dplyr::ungroup()

  gg_post <- tidy_bayes_dat %>%   
    dplyr::select(!!terms, chain) %>% 
    tidyr::gather(term, estimate, -chain) %>% 
    ggplot2::ggplot(aes(estimate, fill = chain, colour = chain)) +
    ggplot2::geom_density(alpha = .3, color = NA) +
    ggplot2::geom_rug(alpha = .1) +
    ggplot2::facet_wrap(~term, ncol = 1, scales = "free_x") +
    ggplot2::theme(legend.position = "bottom")  +
    ggplot2::ggtitle("Marignal Posterior Distrubtion") +
    ggthemes::scale_fill_gdocs()
  
  gg_trace <- tidy_bayes_dat %>%   
    dplyr::select(!!terms, chain, step) %>% 
    tidyr::gather(term, estimate, -chain, -step) %>% 
    ggplot2::ggplot(aes(step, estimate, group = chain, colour = chain)) +
    ggplot2::geom_line(size = .5, alpha = .5) +
    ggplot2::geom_smooth() +
    ggplot2::facet_wrap( ~ term, ncol = 1, scales = "free_y") +
    ggplot2::theme(legend.position = "bottom") +
    ggplot2::ggtitle("Trace Plot of MCMC Chains") +
    ggthemes::scale_colour_gdocs()

  return(gridExtra::grid.arrange(gg_trace, gg_post, ncol = 2))
}
```


## Data {-}

A reduced dataset of Student Panel Survey during the Lecture in Introduction to Political Methodology Winter term 2016/2017 at the University of Konstanz 

* `poleff` Political Efficacy (Likert Score based on 7 items) A larger value = higher level of efficacy
* `friend` Number of alteri in friendship network
* `poldisc` Number of alteri in political discussion network
* `lr.self` Ideological orientation (left right self-placement) 1: Left <- -> 11: Right
* `lr.self.2` Ideological orientation (left right self-placement, second measurement) 1: Left <- -> 11: Right
* `univ.election` Vote intention at the next university election. 1: Yes; 0: other (No and DK)
* `polint` interest at university politics 1: not interested at all <- -> 5 strongly interested
* `tuition` opinion on the general tuition fee for German universities 1: support; 2: reject; 3: indifferent
* `acceptable` acceptable level of the tuition fee (in Euro per Semester) (Only those who support the tuition fee or indifferent)
* protest1 - protest6 willingness to participate a protest action against the general tuition fee 1: yes; 0: no
    + `protest1` demonstration in Konstanz 
    + `protest2` demonstration in Stuttgart 
    + `protest3` giving signature at petitions 
    + `protest4` strike 
    + `protest5` occupation of university buildings 
    + `protest6` legal dispute at courts


```{r}
dat <- get(load("data/Bayes_Student_Survey.RData")) %>% 
  drop_na(univ.election, lr.self)

dat %>% glimpse
```



# Estimate the parameters of a binary logit model. 

You can choose a dependent variable and one independent variable from the dataset for yourself.

```{r binary_model, echo = T}
binary_model1 <- "model{
  for (i in 1:N){
    y[i] ~ dbern(p[i])
    logit(p[i]) <- ystar[i]
    ystar[i] <- alpha + beta * x[i]
  }

  alpha ~ dnorm(0,0.0001)
  beta ~ dnorm(0,0.0001)
}"


write(binary_model1, "bivariate_binary_model1.bug")
```


```{r}
jags.data <- list(
  y = dat$univ.election,
  x = dat$lr.self,
  N = nrow(dat)
)

jags.inits <- 1:5 %>% 
  map(~ list(beta = runif(1, min = 0, max = 1)))

fit_binary_1 <- jags.model(
  file = "bivariate_binary_model1.bug",
  inits = jags.inits,
  data = jags.data, 
  n.chains = length(jags.inits)
)
```


# Calculate DIC of the logit model above by computing the log-likelihood value by using JAGS. 

You can orient yourself the code on the slide 7.18, but you have to care about which distribution the logit model assumes.

```{r binary_model2, echo = T}
binary_model2 <- "model{
  for (i in 1:N){
    y[i] ~ dbern(p[i])
    logit(p[i]) <- ystar[i]
    ystar[i] <- alpha + beta * x[i]
    ### LL for Binary Data - From: https://data.princeton.edu/wws509/notes/c3.pdf
    ll[i] <- y[i]*log(p[i]) + (1 - y[i]) * log(1 - p[i])
  }

  alpha ~ dnorm(0,0.0001)
  beta ~ dnorm(0,0.0001)

  LL <- sum(ll[]) 
}"


write(binary_model2, "bivariate_binary_model2.bug")

fit_binary_2 <- jags.model(
  file = "bivariate_binary_model2.bug",
  inits = jags.inits,
  data = jags.data, 
  n.chains = length(jags.inits)
)

fit_binary_2_list <- coda.samples(
  fit_binary_2,
  variable.names = c("alpha", "beta", "LL"),
  n.iter = 5000,
  thin = 5
)

## Average LL and Deviance
LL <- fit_binary_2_list %>% map(~.x[,"LL"])
LL %>% map_dbl(mean)
Deviances <- LL %>% map(~ .x * -2)
Deviances %>% map_dbl(mean)
```



# Calculate DIC of the logit model above by using dic.samples().

```{r}
binary_dic <- dic.samples(fit_binary_1, n.iter = 2000, thin = 1)
binary_dic

binary_dic_list <- coda.samples(
  fit_binary_1,
  variable.names=c("alpha", "beta", "deviance", "LL"),
  n.iter = 2000, 
  thin = 1
)

# No LL aviable?
# binary_dic_list %>% 
#   map(~mean(.x[, "LL"])) 

binary_dic_list %>% 
  map(~mean(.x[, "deviance"])) 
```


```{r}
plot_jags_model(binary_dic_list, terms = c("alpha", "beta", "deviance"))
```

```{r}
sessionInfo()
```

