---
title: "Lab Session 5"
output: 
# pdf_document
  semantic.doc::semantic_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>
<br>
<div class = "ui text container">

<h1 class = "ui header">
 Lab Session 5
</h1>


* posterior correlated := autokorrelation or multicolinearity.


# Packages 

```{r}
pacman::p_load(rjags, dplyr, purrr, tidyr, ggplot2, broom, rjags)
ggplot2::theme_set(theme_bw())
# tinytex::install_tinytex()
devtools::session_info()
set.seed(2018)
rjags::
```



# Bayes Regression: Beta Binomial

## Data

* `p` = probability of success
* `N` = Number of trials

```{r}
a <- 1 
b <- 1

beta_dist <- function(x) dbeta(x, a, b)
curve(beta_dist, 0, 1)
```



```{r}
binom.model <- "model{
   y~dbin(p, N)
   p ~ dbeta(a, b) # Prior
}"

write(binom.model, "Bayes_Binom_Beta.bug")

# Data 
jags.data <- list(y = 60, N = 100, a = 3, b = 1.5)

# Running JAGS
jags.reg <- jags.model(
  file = "Bayes_Binom_Beta.bug",
  data = jags.data, 
  n.chains = 3
)

# Burn in phase (warm up)
update(jags.reg, 1000)

jags.out <- coda.samples(
  jags.reg,
  variable.name =  c("p"),
  n.iter = 1000, 
  thin = 1
)

# Simple description of posterior
summary(jags.out)
plot(jags.out)

# Which percentage of posterior p>0.5 ?
p <- unlist(jags.out)

tibble(dens = p > 0.5) %>% 
  janitor::tabyl(dens)

hist(p)
plot(density(p))
```


# Bayes Regression: Student Survey

## Data

```{r}
dat <- get(load("data/Bayes_Student_Survey.RData")) %>% 
  mutate(friend_log = log(friend + 1))

glimpse(dat)
```


A reduced dataset of Student Panel Survey during the Lecture in Introduction to Political Methodology Winter term 2016/2017 at the University of Konstanz 

* `poleff` Political Efficacy (Likert Score based on 7 items) A larger value = higher level of efficacy
* `friend` Number of alteri in friendship network
* `poldisc` Number of alteri in political discussion network
* `lr.self` Ideological orientation (left right self-placement) 1: Left <- -> 11: Right
* `lr.self.2` Ideological orientation (left right self-placement, second measurement) 1: Left <- -> 11: Right
* `univ.election` Vote intention at the next university election. 1: Yes; 0: other (No and DK)
* `polint` interest at university politics 1: not interested at all <- -> 5 strongly interested
* `tuition` opinion on the general tuition fee for German universities 1: support; 2: reject; 3: indifferent
* `acceptable` acceptable level of the tuition fee (in Euro per Semester) (Only those who support the tuition fee or indifferent)
* protest1 - protest6 willingness to participate a protest action against the general tuition fee 1: yes; 0: no
    + `protest1` demonstration in Konstanz 
    + `protest2` demonstration in Stuttgart 
    + `protest3` giving signature at petitions 
    + `protest4` strike 
    + `protest5` occupation of university buildings 
    + `protest6` legal dispute at courts


> **Task:** Estimate the effect of the size of friendship network on pol. efficacy


## OLS regression

```{r}
dat %>% 
  select(poleff, friend) %>% 
  gather(varname, value) %>% 
  ggplot(aes(value)) +
  geom_bar(alpha = .6) +
  facet_wrap(~varname, scales = "free_x")
```


```{r}
dat %>%
  ggplot(aes(friend_log, poleff)) +
  geom_jitter()
```


```{r}
ols.out <- lm(poleff ~ friend_log, data=dat)
summary(ols.out)
```

```{r}
ols.out %>% 
  broom::augment() %>%
  ggplot(aes(friend_log, poleff)) +
  geom_jitter() +
  geom_line(aes(friend_log, .fitted))
  #geom_smooth(method = "lm")
```


## rjags


$$y_i \sim N(\mu, \tau)$$
$$\mu = \beta_0 + \beta_1 x_1$$


* tau != variance but precision (inverse of variance); with certain tau we have certain sigma, not specified in then model. 
* what do with sigma? is determined by tau, only for extractiuon purpose.


```{r}
# JAGS Modell
reg.model <- "model{
  for (i in 1:N){
    y[i] ~ dnorm(mu[i], tau)
    mu[i] <- beta0 + beta1 * x[i]
  }
  
  beta0 ~ dnorm(0, 0.0001)
  beta1 ~ dnorm(0, 0.0001) 
  
  tau ~ dgamma(0.001, 0.001) 
  sigma <- 1/sqrt(tau)
}"

write(reg.model, "Bayes_Bivariate_Reg_Student_Survey.bug")
```


```{r}
jags.data <- list(
  y = dat$poleff,
  x = dat$friend_log,
  N = length(dat$poleff)
)

# three different intial values for beta1
jags.inits <- list(
  list(beta1 = 323),
  list(beta1 = 5000),
  list(beta1 = -10)
)

# jags_init <- function(){
# list(
#   list(beta1 = 323),
#   list(beta1 = 5000),
#   list(beta1 = -10)
# )
# }
# jags_init()

jags.reg <- jags.model(
  file = "Bayes_Bivariate_Reg_Student_Survey.bug",
  inits = jags.inits,
  data = jags.data, 
  n.chains=3
)
```

```{r}
update(jags.reg, 2000)
jags.reg.out <- coda.samples(jags.reg,
                             variable.names=c("beta0","beta1","sigma"),
                             n.iter=2000, thin=1)

jags.reg.out %>% 
  summary() %>% 
  .$statistics
```

* Question: Do we have autokorrelation in the data? Why?
* Does the SE gets smaller through thinnin?
* SD of posterior can be large if inital values are included.
* 

## Posterior Estimates

```{r}
tidy_bayes_dat <- jags.reg.out %>% 
  tibble(output = .) %>% 
  mutate(chain = as.character(1:n())) %>%
  mutate(output = map(output, as_tibble)) %>% 
  tidyr::unnest() %>% 
  group_by(chain) %>% 
  mutate(step = 1:n()) %>% 
  ungroup
```


```{r}
(gg_post <- tidy_bayes_dat %>%   
  select(beta0, beta1, sigma, chain) %>% 
  gather(term, estimate, -chain) %>% 
  ggplot(aes(estimate, fill = chain, colour = chain)) +
  geom_density(alpha = .3, color = NA) +
  # geom_eyeh() + 
  #geom_vline(xintercept = mean(tidy_bayes_dat$sigma), linetype = 2) +
  geom_rug(alpha = .1) +
  facet_wrap(~term, ncol = 1, scales = "free_x") +
  theme(legend.position = "bottom")  +
  ggtitle("Marignal Posterior Distrubtion"))
```

```{r}
tidy_bayes_dat %>%   
  select(beta0, beta1, chain) %>% 
  gather(term, estimate, -chain) %>% 
  group_by(term) %>%
  summarise(estimate = list(quantile(estimate, ppoints(100)))) %>% 
  ungroup %>% 
  unnest %>% 
  ggplot(aes(x = estimate)) +
  geom_dotplot(binwidth = .04) +
  facet_wrap(~term, ncol = 1, scales = "free_x")
```




```{r}
(gg_trace <- tidy_bayes_dat %>%   
  select(beta0, beta1, sigma, chain, step) %>% 
  gather(term, estimate, -chain, -step) %>% 
  ggplot(aes(step, estimate, group = chain, colour = chain)) +
  geom_line(size = .5, alpha = .5) +
  facet_wrap(~term, ncol = 1, scales = "free_y") +
  theme(legend.position = "bottom") +
  ggtitle("Trace Plot of MCMC Chains"))
```


```{r}
gridExtra::grid.arrange(gg_trace, gg_post, ncol = 2)
```



</div>