---
title: "Bayesian Statistics"
subtitle: "Class Assignment 1"
author: "Simon Roth \\\n nomis.roth@gmx.net"
date: "28.11.2018"
output: 
  pdf_document:
    number_sections: true
---


```{r setup, include=F}
knitr::opts_chunk$set(echo = T, fig.height = 3, fig.width = 4, fig.align = "center")
```


## Packages {-}

**Note** *This is Rmarkdown document which combines the entire code, outputs and text. If you don't have pacman installed, just do it once and it will manage all the rest of the dependecies (for ever).*

```{r}
pacman::p_load(rjags, dplyr, purrr, tidyr, ggplot2, broom, rjags, texreg, ggthemes, janitor, knitr)
ggplot2::theme_set(theme_bw())
set.seed(2018)
```


## Data {-}

A reduced dataset of Student Panel Survey during the Lecture in Introduction to Political Methodology Winter term 2016/2017 at the University of Konstanz 

* `poleff` Political Efficacy (Likert Score based on 7 items) A larger value = higher level of efficacy
* `friend` Number of alteri in friendship network
* `poldisc` Number of alteri in political discussion network
* `lr.self` Ideological orientation (left right self-placement) 1: Left <- -> 11: Right
* `lr.self.2` Ideological orientation (left right self-placement, second measurement) 1: Left <- -> 11: Right
* `univ.election` Vote intention at the next university election. 1: Yes; 0: other (No and DK)
* `polint` interest at university politics 1: not interested at all <- -> 5 strongly interested
* `tuition` opinion on the general tuition fee for German universities 1: support; 2: reject; 3: indifferent
* `acceptable` acceptable level of the tuition fee (in Euro per Semester) (Only those who support the tuition fee or indifferent)
* protest1 - protest6 willingness to participate a protest action against the general tuition fee 1: yes; 0: no
    + `protest1` demonstration in Konstanz 
    + `protest2` demonstration in Stuttgart 
    + `protest3` giving signature at petitions 
    + `protest4` strike 
    + `protest5` occupation of university buildings 
    + `protest6` legal dispute at courts


```{r}
dat <- get(load("data/Bayes_Student_Survey.RData")) %>% 
  mutate(friend_log = log(friend + 1))

# glimpse(dat)
```


# Frequentist Estimation

Estimate the parameters of a bivariate regression via OLS. You can choose a dependent variable and one independent variable from the dataset for yourself.

Political interest might influence the general openess to communicate and make friends. 


```{r, results="asis"}
dat %>% 
  ggplot(aes(polint, friend_log)) +
  geom_jitter() +
  geom_smooth(method = "lm")

lm(friend_log ~ polint, data = dat) %>% 
  texreg::texreg(float.pos = "ht!")
```


# Bayesian Estimation

Run the MCMC to obtain the posterior of the same regression model above with 5 chains. You have to run the first 200 iterations without collecting posterior. Thereafter collect your posterior in 1000 iterations. Use the same prior as on the slides. 


```{r jagsmodel1, echo = T}
reg.model <- "model{
  for (i in 1:N){
    y[i] ~ dnorm(mu[i], tau)
    mu[i] <- beta0 + beta1 * x[i]
  }
  
  beta0 ~ dnorm(0, 0.0001)
  beta1 ~ dnorm(0, 0.0001) 
  
  tau ~ dgamma(0.001, 0.001) 
  sigma <- 1/sqrt(tau)
}"

write(reg.model, "Bayes_Bivariate_Reg_Student_Survey.bug")
```


```{r}
jags.data <- list(
  y = dat$friend_log,
  x = dat$polint,
  N = nrow(dat)
)

jags.inits <- 1:5 %>% 
  map(~ list(beta1 = runif(1, min = -100, max = 100) %>% round()))

jags.reg <- jags.model(
  file = "Bayes_Bivariate_Reg_Student_Survey.bug",
  inits = jags.inits,
  data = jags.data, 
  n.chains = length(jags.inits)
)
```


```{r}
update(jags.reg, 200)

jags.reg.out <- coda.samples(
  jags.reg,
  variable.names = c("beta0", "beta1", "sigma"),
  n.iter = 1000, 
  thin = 1
)

jags.reg.out %>% 
  summary() %>% 
  .$statistics %>% 
  kable
```

# Check Cnvergence 

based on visible inspection and the Gelman- Rubin-Statistics.

```{r, fig.width = 8, fig.height = 6}
gelman.plot(jags.reg.out)
#autocorr.plot(jags.reg.out)
```

# Report the posterior 

by using summary() and plot().

```{r, fig.width = 8, fig.height = 6, echo = F}
plot_jags_model <- function(model, terms = c("beta0", "beta1", "sigma")){
  
  tidy_bayes_dat <- model %>% 
    dplyr::tibble(output = .) %>% 
    dplyr::mutate(chain = as.character(1:n())) %>%
    dplyr::mutate(output = purrr::map(output, as_tibble)) %>% 
    tidyr::unnest() %>% 
    dplyr::group_by(chain) %>% 
    dplyr::mutate(step = 1:n()) %>% 
    dplyr::ungroup()

  gg_post <- tidy_bayes_dat %>%   
    dplyr::select(!!terms, chain) %>% 
    tidyr::gather(term, estimate, -chain) %>% 
    ggplot2::ggplot(aes(estimate, fill = chain, colour = chain)) +
    ggplot2::geom_density(alpha = .3, color = NA) +
    ggplot2::geom_rug(alpha = .1) +
    ggplot2::facet_wrap(~term, ncol = 1, scales = "free_x") +
    ggplot2::theme(legend.position = "bottom")  +
    ggplot2::ggtitle("Marignal Posterior Distrubtion") +
    ggthemes::scale_fill_hc()
  
  gg_trace <- tidy_bayes_dat %>%   
    dplyr::select(!!terms, chain, step) %>% 
    tidyr::gather(term, estimate, -chain, -step) %>% 
    ggplot2::ggplot(aes(step, estimate, group = chain, colour = chain)) +
    ggplot2::geom_line(size = .5, alpha = .5) +
    ggplot2::geom_smooth() +
    ggplot2::facet_wrap(~term, ncol = 1, scales = "free_y") +
    ggplot2::theme(legend.position = "bottom") +
    ggplot2::ggtitle("Trace Plot of MCMC Chains") +
    ggthemes::scale_colour_hc()

  return(gridExtra::grid.arrange(gg_trace, gg_post, ncol = 2))
}

plot_jags_model(jags.reg.out)
```

# Calculate the probability that beta1 is positive

Which percentage of posterior is greater than zero (positive)?

```{r}
unlist(jags.reg.out) %>% 
  tibble(p = . > 0) %>% 
  tabyl(p) %>% 
  kable
```

Answer: 99% 


# Repeat 2-4 with different prior.

